---
title: "Chapter 2 HW"
author: "Fabiani Rafael"
#date: "Due date of Chapter 2 HW"
output:
  html_document:
    fig_height: 4
    highlight: pygments
    theme: spacelab
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

* * *

## Conceptual Questions


#### Exercise 2: Explain whether each scenario is a classification or regression prob-lem, and indicate whether we are most interested in inference or prediction. Finally, provide n := sample size and p := predictors.

(a) We collect a set of data on the top 500 firms in the US. For each firm we record profit, number of employees, industry and the CEO salary. We are interested in understanding which factors affect CEO salary.

- This describes an instance of regressional inference because our y i.e the CEO salary is a continous variable and it is an example of an inference problem since the goal is not primarily to predict a given output but more so the decipher the underlying relationships between profit, number of employees, industry and the CEO salary. Regression fits this use case because instead of primarily focusing upon using input to predict changes in the CEO salary, we want to see which of these factors effect CEO salary not necessarily how they will effect CEO salary. Here we have $n = 500 \, , \, p = 3$.

(b) We are considering launching a new product and wish to know whether it will be a success or a failure. We collect data on 20 similar products that were previously launched. For each product we have recorded whether it was a success or failure, price charged for the product, marketing budget, competition price, and ten other variables.

- This describes an instance of a classification problem geared towards prediction. This falls into the category of classification as it deals with a y (success/failuere) which is categorical in nature moreover its goal is to predict the success or failure of the new product using collected data from similar products. Here $n = 20 \, , \, p = 13$


(c) We are interested in predicting the % change in the USD/Euro exchange rate in relation to the weekly changes in the world stock markets. Hence we collect weekly data for all of 2012. For each week we record the % change in the USD/Euro, the % change in the US market, the % change in the British market, and the % change in the German market.

- This is an example of a regressional prediction problem. It falls into the category of regression as it deals with a y which is continous in nature and fits into the category of prediction since its primary goal is to use input data i.e the precent change in the US, British and German markets to predict the % change in the USD/Euro exchange rate. Here $n=52 \, , \, p = 3$.

#### Exercise 4: You will now think of some real-life applications for statistical learn-ing.
(a) Describe three real-life applications in which classification might be useful. Describe the response, as well as the predictors. Is the goal of each application inference or prediction? Explain your answer.

- One example of a real life application in which classification would be useful is in sentiment analysis possibly for something like product reviews. Where in the response could be a sentiment label i.e an indicator for Positive, Negative and possibly Neutral reception. Then some predictors could be Text data ( for isntance review  wording), star ratings, and reviewer history to gauge tendency of a user to leave either highly negative or positive reviews. The goal of this example would be primarily prediction since our intent is to classiffy new product reviews into categories based on learned patterns from data.The model would be trained on labeled reviews where the sentiment is already known and then it would predict the sentiment of unseen reviews.

- Another example of a real life application could be Handwritten digit recognition where in the response would be a digit label i.e (0-9) and the predictors could be pixel valus of the image, edge detection features and shape descriptors. The primary goal would be prediction i.e trying to predict the correct digit for new/unseen images. So the model would take an input of a handwritten digit and classify it as one of the digits 0-9 and then use that to predit the most likely represents.

- A real life application could be in transaction classification i.e fraudulent or legitimate. Predictors such as transaction amount, time of the transaction , location , device used and transaction history. The goal here would again be precition becuase the main intent would be to identify fraudulent transactions in real time using past transaction data to train out model with transactions labeled as either fraudulent or legitimate. When a new transaction is made the model would predict whether it is a valid transaction i.e non fraudulent based on its feautures. So the task here would be classifying new transactions rather than undersrantinng the causes of fraud.

(b) Describe three real-life applications in which regression might be useful. Describe the response, as well as the predictors. Is the goal of each application inference or prediction? Explain your answer.

- One real-life application where regression may be found useful is in the analysis of drivers of life expectancy. A response value for this would be the persons life expectancy at birth (a continous val. in years). Predictors in this example could be things like GDP per capita, access to healthcare, sanitation, education, and other social factors. The goal of this application would be inference since the main intent would be to understand the relationship between the predictors and the response i.e life expectancy. The model would be used to understand how the predictors effect life expectancy and not necessarily to predict life expectancy.

- Another example of an application where regression would be of use is in forcasting retail demand. A response value for this use case could be Daily sales volume (units sold) and the predictors could be things like day of the week, time of the year, weather, promotions, competitor pricing and maybe historical data like sales from the past 30 days. The goal of this application would be prediction since the main intent would be to predict future sales volume based on past sales data and other predictors to gauge future demand for inventory management. The model would be used to predict future sales volume based on past sales data and could help in anticipating holiday sales for a specific product to optimize stock levels.

- An example of an application that could use regression is also in predicting renewable energy production i.e solar power, wind farms. A response value for this could be the amount of energy produced in a given time period. Predictors could be things like weather data for instance wind speed, temperature and air pressure, time of day, season, and location. The goal of this application would be prediction since the main intent would be to predict the amount of energy produced in the future based on past data. The model would be used to predict future energy production based on trends in previous energy production and could help in optimizing energy storage and distribution.

(c) Describe three real-life applications in which cluster analysis might be useful.

- One real-life application where cluster analysis might be useful is in customer segmentation. For instance in a retail setting, where the goal would be to segment customers into distinct groups based on their purchasing behavior / history. Clustered data could be customer transaction history , customer demographics, and engagement metrics. Predictors could be things like purchase frequency, average amount spent per order and product preferences. Demographic data that could be used can be age, location, and income and behaviral data would be things like website click or email open rates. This could help in tailoring marketing strategies to each group. For instance, a cluster of customers who buy mostly electronics could be targeted with electronics promotions because they may be more likely to be interested in those products.

- Another real-life application where cluster analysis might be useful is in anomaly detection. For instance in network security where the goal would be to detect unusual behavior in network traffic. Clustered data could be network traffic data, user activity logs, and system logs such access attempts. Features here would be things like data transfer rates, login attempts, IP addresses, file access times, protocol type and time of day. This could help in identifying unusual patterns in network traffic that could be indicative of a security breach becuase clustering may help in distinguishing normal activity from anomalous clusters and security teams could prioritize investigating outliers flagged by the model.

- Another real-life application where cluster analysis might be useful is in recommendation systems. For instance in a streaming service like Netflix where the goal would be to recommend movies or shows to users based on their viewing history. Clustered data could be user behavior for isntance viewing history, user ratings, search queries, and time spent on content or potentially content attributes like genre, actors, directors, and release year. Features here could be viewing habits, with similar themes or audience appeal as content clusters. This could help in recommending movies or shows to users based on their viewing history and preferences. For instance, a user who watches a lot of action movies could be recommended more action movies or a user who watches a lot of comedies could be recommended more comedies.


* * *

## Applied Questions

#### Exercise 8: This exercise relates to the College data set, which can be found in the file College.csv on the book website. It contains a number of variables for 777 different universities and colleges in the US. The variables are

• Private : Public/private indicator

• Apps : Number of applications received

• Accept : Number of applicants accepted

• Enroll : Number of new students enrolled

• Top10perc : New students from top 10 % of high school class

• Top25perc : New students from top 25 % of high school class

• F.Undergrad : Number of full-time undergraduates

• P.Undergrad : Number of part-time undergraduates

• Outstate : Out-of-state tuition

• Room.Board : Room and board costs

• Books : Estimated book costs

• Personal : Estimated personal spending

• PhD : Percent of faculty with Ph.D.’s

• Terminal : Percent of faculty with terminal degree

• S.F.Ratio : Student/faculty ratio

• perc.alumni : Percent of alumni who donate

• Expend : Instructional expenditure per student

• Grad.Rate : Graduation rate

Before reading the data into R, it can be viewed in Excel or a 
text editor.

(a) Use the read.csv() function to read the data into R. Call the loaded data college. Make sure that you have the directory set to the correct location for the data.

``` {r}
college <- read.csv("College.csv")
```

(b) Look at the data using the View() function. You should notice that the first column is just the name of each university. We don’t really want R to treat this as data. However, it may be handy to have these names for later. Try the following commands:

> rownames(college) <- college[, 1]

> View(college)

You should see that there is now a row.names column with the name of each university recorded. This means that R has given each row a name corresponding to the appropriate university. R will not try to perform calculations on the row names. However, we still need to eliminate the first column in the data where the names are stored. Try

> college <- college[, -1]

> View(college)

Now you should see that the first data column is Private. Note that another column labeled row.names now appears before the Private column. However, this is not a data column but rather the name that R is giving to each row.

(c) 
i. Use the summary() function to produce a numerical summary of the variables in the data set.56

ii. Use the pairs() function to produce a scatterplot matrix of the first ten columns or variables of the data. Recall that you can reference the first ten columns of a matrix A using  A[,1:10].

iii. Use the plot() function to produce side-by-side boxplots of Outstate versus Private.

iv. Create a new qualitative variable, called Elite, by binning the Top10perc variable. We are going to divide universities into two groups based on whether or not the proportion of students coming from the top 10 % of their high school classes exceeds 50 %.

> Elite <- rep("No", nrow(college))

> Elite[college$Top10perc > 50] <- "Yes"

> Elite <- as.factor(Elite)

> college <- data.frame(college , Elite)

Use the summary() function to see how many elite universities there are. Now use the plot() function to produce side-by-side boxplots of Outstate versus Elite. v. Use the hist() function to produce some histograms with differing numbers of bins for a few of the quantitative variables. You may find the command par(mfrow = c(2, 2)) useful: it will divide the print window into four regions so that four plots can be made simultaneously. Modifying the arguments to this function will divide the screen in other ways.

vi. Continue exploring the data, and provide a brief summary of what you discover.

### Exercise 10 This exercise involves the Boston housing data set.
(a) To begin, load in the Boston data set. The Boston data set is part of the ISLR2 library.

> library(ISLR2)

Now the data set is contained in the object Boston.

> Boston

Read about the data set:

> ?Boston

How many rows are in this data set? How many columns? What
do the rows and columns represent?

(b) Make some pairwise scatterplots of the predictors (columns) in this data set. Describe your findings.

(c) Are any of the predictors associated with per capita crime rate? If so, explain the relationship.

(d) Do any of the census tracts of Boston appear to have particularly high crime rates? Tax rates? Pupil-teacher ratios? Comment on the range of each predictor.

(e) How many of the census tracts in this data set bound the Charles river?

(f) What is the median pupil-teacher ratio among the towns in this data set?

(g) Which census tract of Boston has lowest median value of owner-occupied homes? What are the values of the other predictors for that census tract, and how do those values compare to the overall ranges for those predictors? Comment on your findings.

(h) In this data set, how many of the census tracts average more than seven rooms per dwelling? More than eight rooms per dwelling? Comment on the census tracts that average more than eight rooms per dwelling. 






