---
title: "Chapter 2 HW"
author: "Fabiani Rafael"
#date: "Due date of Chapter 2 HW"
output:
  pdf_document: default
  html_document:
    fig_height: 4
    highlight: pygments
    theme: spacelab
  word_document: default
---

```{r setup, include=FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

* * *

## Conceptual Questions


#### Exercise 2: Explain whether each scenario is a classification or regression prob-lem, and indicate whether we are most interested in inference or pre-diction. Finally, provide n and p.

(a) We collect a set of data on the top 500 firms in the US. For each
firm we record profit, number of employees, industry and the
CEO salary. We are interested in understanding which factors
affect CEO salary.
(b) We are considering launching a new product and wish to know
whether it will be a success or a failure. We collect data on 20
similar products that were previously launched. For each prod-
uct we have recorded whether it was a success or failure, price
charged for the product, marketing budget, competition price,
and ten other variables.
(c) We are interested in predicting the % change in the USD/Euro
exchange rate in relation to the weekly changes in the world
stock markets. Hence we collect weekly data for all of 2012. For
each week we record the % change in the USD/Euro, the %
change in the US market, the % change in the British market,
and the % change in the German market.

#### Exercise 4: You will now think of some real-life applications for statistical learn-ing.
(a) Describe three real-life applications in which classification might
be useful. Describe the response, as well as the predictors. Is the
goal of each application inference or prediction? Explain your
answer.
(b) Describe three real-life applications in which regression might
be useful. Describe the response, as well as the predictors. Is the
goal of each application inference or prediction? Explain your
answer.
(c) Describe three real-life applications in which cluster analysis
might be useful.

* * *

## Applied Questions

#### Exercise 8: This exercise relates to the College data set, which can be found in
the file College.csv on the book website. It contains a number of
variables for 777 different universities and colleges in the US. The
variables are
• Private : Public/private indicator
• Apps : Number of applications received
• Accept : Number of applicants accepted
• Enroll : Number of new students enrolled
• Top10perc : New students from top 10 % of high school class
• Top25perc : New students from top 25 % of high school class2.4 Exercises
55
• F.Undergrad : Number of full-time undergraduates
• P.Undergrad : Number of part-time undergraduates
• Outstate : Out-of-state tuition
• Room.Board : Room and board costs
• Books : Estimated book costs
• Personal : Estimated personal spending
• PhD : Percent of faculty with Ph.D.’s
• Terminal : Percent of faculty with terminal degree
• S.F.Ratio : Student/faculty ratio
• perc.alumni : Percent of alumni who donate
• Expend : Instructional expenditure per student
• Grad.Rate : Graduation rate
Before reading the data into R, it can be viewed in Excel or a text
editor.
(a) Use the read.csv() function to read the data into R. Call the
loaded data college. Make sure that you have the directory set
to the correct location for the data.
(b) Look at the data using the View() function. You should notice
that the first column is just the name of each university. We don’t
really want R to treat this as data. However, it may be handy to
have these names for later. Try the following commands:

> rownames(college) <- college[, 1]
> View(college)

You should see that there is now a row.names column with the
name of each university recorded. This means that R has given
each row a name corresponding to the appropriate university. R
will not try to perform calculations on the row names. However,
we still need to eliminate the first column in the data where the
names are stored. Try
> college <- college[, -1]
> View(college)
Now you should see that the first data column is Private. Note
that another column labeled row.names now appears before the
Private column. However, this is not a data column but rather
the name that R is giving to each row.
(c)
i. Use the summary() function to produce a numerical summary
of the variables in the data set.56
2. Statistical Learning
ii. Use the pairs() function to produce a scatterplot matrix of
the first ten columns or variables of the data. Recall that
you can reference the first ten columns of a matrix A using
A[,1:10].
iii. Use the plot() function to produce side-by-side boxplots of
Outstate versus Private.
iv. Create a new qualitative variable, called Elite, by binning
the Top10perc variable. We are going to divide universities
into two groups based on whether or not the proportion
of students coming from the top 10 % of their high school
classes exceeds 50 %.
> Elite <- rep("No", nrow(college))
> Elite[college$Top10perc > 50] <- "Yes"
> Elite <- as.factor(Elite)
> college <- data.frame(college , Elite)
Use the summary() function to see how many elite univer-
sities there are. Now use the plot() function to produce
side-by-side boxplots of Outstate versus Elite.
v. Use the hist() function to produce some histograms with
differing numbers of bins for a few of the quantitative vari-
ables. You may find the command par(mfrow = c(2, 2))
useful: it will divide the print window into four regions so
that four plots can be made simultaneously. Modifying the
arguments to this function will divide the screen in other
ways.
vi. Continue exploring the data, and provide a brief summary
of what you discover.

### Exercise 10 This exercise involves the Boston housing data set.
(a) To begin, load in the Boston data set. The Boston data set is
part of the ISLR2 library.
> library(ISLR2)
Now the data set is contained in the object Boston.
> Boston
Read about the data set:
> ?Boston
How many rows are in this data set? How many columns? What
do the rows and columns represent?
(b) Make some pairwise scatterplots of the predictors (columns) in
this data set. Describe your findings.
(c) Are any of the predictors associated with per capita crime rate?
If so, explain the relationship.
(d) Do any of the census tracts of Boston appear to have particularly
high crime rates? Tax rates? Pupil-teacher ratios? Comment on
the range of each predictor.
(e) How many of the census tracts in this data set bound the Charles
river?
(f) What is the median pupil-teacher ratio among the towns in this
data set?
(g) Which census tract of Boston has lowest median value of owner-
occupied homes? What are the values of the other predictors
for that census tract, and how do those values compare to the
overall ranges for those predictors? Comment on your findings.
(h) In this data set, how many of the census tracts average more than
seven rooms per dwelling? More than eight rooms per dwelling?
Comment on the census tracts that average more than eight
rooms per dwelling. 

```{r}
summary(cars)
plot(cars)
```







